### 2015-08-03 chapters/conclusion.tex
added intro and conclusion outlines

\chapter{Conclusions} \label{chapter:conclusion}

Summary of previous chapters here.

\section{Recent Developments}

Go through some recent/concurrent papers that may have built on these ideas.

Resolving magnetic fields: \citep{stephens14}

HL Tau: \citep{brogan14}

A gap in Tw Hya at 20 AU? \citep{akiyama15}

Gas density within cavities of transition disks \citep{vandermarel15}

More dust traps: \citep{espaillat15}

\section{A Bright Future}
Circumplanetary disks? \citep{perez15}
Chemistry and stuff \citep{oberg15} \citep{graninger15} 
Chemistry as a signpost of planet formation \citep{cleeves15}
Exploring the outer solar system with New Horizons
Detection of organics by Rosetta? Checkout: \url{http://www.sciencedaily.com/releases/2015/07/150730172518.htm}

Summarize status of EHT and its future.