### 2015-08-03 chapters/conclusion.tex
chapter 2 summary

\chapter{Conclusions} \label{chap:conclusion}

One might argue that the space of discovery in Astronomy is spanned by resolution and 
wavelength.  Technology has enabled us to contemplate phenomena at wavelengths far outside 
what the realm of our human senses, revolutionizing fields like high energy astrophysics,
cosmology, and the study of the inter stellar-medium.  Resolution then reveals the hidden
details and helps us to interpret what these new eyes see.  A recent example is 
the application of interferometry at sub-mm wavelengths with ground breaking telescopes like the SMA, 
the amazing new power of ALMA, and the EHT revolution.  
While the science motivating chapters \ref{chap:v4046_structure} 
through \ref{chap:radial_flows} may be different than in the final chapter of this thesis, 
I am left wondering and waiting for what will we see when we look only a little bit closer.

Emboldened by this spirit, we shall conclude not only with summaries of the work presented here,
but also some brief discussion of recent developments in the field.  In chapter \ref{chap:v4046_structure} we 
used the SMA to study the protoplanetary disk 
around the close binary V4046 Sgr.  This disk featured a narrow ring of large mm-sized grains, 
but an extended gas rich disk.  Detailed modeling of these features demonstrated that they were 
phenomenologically consistent with the viscous spreading of the gas and the concentration through pressure 
drag forces of the dust.  This disk is a single example among a number, and more such examples continue to be 
found, \citep{isella07,panic09,andrews12,degregorio-monsalvo13,ke14}.  A somewhat related phenomenon are the 
severely asymmetric dust  ``horse-shoes'' that have been recently observed at multiple wavelengths
\citep{vandermarel13,fukagawa13,isella13,casassus15,marino15}.  Thought to be perhaps generated by gigantic 
3D Rossby wave instabilities \citep{regaly13}, these are another striking example from the interplay between
the gas and dust in protoplanetary disks.

Resolving magnetic fields: \citep{stephens14}

HL Tau: \citep{brogan14}

A gap in Tw Hya at 20 AU? \citep{akiyama15}

Gas density within cavities of transition disks \citep{vandermarel15}

More dust traps: \citep{espaillat15}

\section{A Bright Future}
What is happening in the inner disks? \citep{ke14}
Circumplanetary disks? \citep{perez15}
Chemistry and stuff \citep{oberg15} \citep{graninger15} 
Chemistry as a signpost of planet formation \citep{cleeves15}
Exploring the outer solar system with New Horizons
Detection of organics by Rosetta? Checkout: \url{http://www.sciencedaily.com/releases/2015/07/150730172518.htm}

Summarize status of EHT and its future.

This thesis is a small expression of 
the science that resolution at these wavelengths can provide and only an exploration for the 
hard and promising work ahead.