### 2015-08-31 frontmatter/authorlist.tex

{\large
\linespread{1.}\selectfont{}
\noindent The following authors contributed to Chapter 1: \newline
Sean M. Andrews, David J. Wilner, J. H. Kastner, M. K. McClure.

\noindent The following authors contributed to Chapter 2: \newline
Sean M. Andrews, A. Meredith Hughes, David J. Wilner.

\noindent The following authors contributed to Chapter 3: \newline
Eugene Chiang and Sean M. Andrews.

\noindent The following authors contributed to Chapter 4: \newline
Lindy Blackburn, Rurik Primiani, Laura Vertatschitsch, and Andre Young.
}