### 2015-09-03 frontmatter/authorlist.tex

{\large
\linespread{1.}\selectfont{}
\noindent The following authors contributed to Chapter 1: \newline
Sean M. Andrews, David J. Wilner, J. H. Kastner, M. K. McClure.

\noindent The following authors contributed to Chapter 2: \newline
Sean M. Andrews, A. Meredith Hughes, David J. Wilner.

\noindent The following authors contributed to Chapter 3: \newline
Eugene Chiang and Sean M. Andrews.

\noindent The following people contributed to Chapter 4: \newline
Lindy Blackburn, Geoff Crew, Sheperd Doeleman, Rurik Primiani, Mike Titus, Laura Vertatschitsch, 
Jonathan Weintroub, and Andre Young
}